---
title: "Time conversion"
author: "Chi Zhang"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Time conversion}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We create the following functions to make it easier to work with date and times in different formats. 


For the input, we allow either a numeric (3) or string ("03", "2020-03"). 

For the output, the user need to specify whether numeric or string is required. Functions end with `c` returns a character output and `n` returns a double. 


```{r}
library(splstyle)
```


## Which year/week is it?

By default the functions returns the isoyear or isoweek of **today**. 

```{r}
isoyear_c()
isoyear_n()
isoweek_c()
isoweek_n()

# provide a date
isoyear_c('2021-01-01')
isoyear_n('2021-01-01')
isoweek_c('2021-01-01')
isoweek_n('2021-01-01')
```




## Get year/week from a `isoyearweek` string

```{r}
isoyearweek_to_year_c("2021-02")
isoyearweek_to_year_n("2021-02")
isoyearweek_to_week_c("2021-02")
isoyearweek_to_week_n("2021-02")
```


A list of `isoyearweek` can be accessed in the following way. 

```{r}
yrwk_19_20 <- fhidata::world_dates_isoyearweek[year %in% c(2019, 2020)]
head(yrwk_19_20)
```




